---
title: "EDA"
output: html_notebook
---

Import Data
```{r}
library(haven)

data <- read_dta("../VisualStudio/yura_stata/fulltrial.dta")
```


Creates comprehensive summary statistics for each cost component by study arm
Generates both box plots and bar plots with error bars
Performs both linear and quantile regression to assess factors associated with total cost
Analyzes and visualizes the proportion each cost component contributes to total cost
Uses efficient data manipulation with tidyverse
Includes proper handling of missing values
Provides appropriate scaling and labeling of visualizations
```{r}
# Load required libraries
library(tidyverse)
library(gtsummary)
library(ggplot2)
library(scales)
library(broom)
library(quantreg)

selected_cols <- c("sv1_twowaycost", "sv1_childcarecost",
                   "sv1_additionalfoodcost", "calculatedwageloss", "p2cost",
                   "sv1_anythingelsecost", "loss_v2")

# Create summary statistics by study arm
cost_summary <- data %>%
  group_by(outcome_twoarm) %>%
  summarise(across(
    selected_cols,
    list(
      median = ~median(., na.rm = TRUE),
      mean = ~mean(., na.rm = TRUE),
      sd = ~sd(., na.rm = TRUE),
      q25 = ~quantile(., 0.25, na.rm = TRUE),
      q75 = ~quantile(., 0.75, na.rm = TRUE),
      n = ~sum(!is.na(.))
    )
  )) %>%
  pivot_longer(cols = -outcome_twoarm,
               names_to = c("cost_type", "stat"),
               names_pattern = "(.+)_(.+)$")

# Create visualizations
# 1. Box plots
cost_boxplot <- data %>%
  pivot_longer(cols = selected_cols,
               names_to = "cost_type",
               values_to = "cost") %>%
  ggplot(aes(x = factor(outcome_twoarm), y = cost, 
             fill = factor(outcome_twoarm))) +
  geom_boxplot() +
  facet_wrap(~cost_type, scales = "free_y") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  labs(x = "Study Arm",
       y = "Cost (USh)",
       fill = "Study Arm") +
  theme(legend.position = "bottom")
plot(cost_boxplot)


# 2. Bar plots with error bars
cost_barplot <- data %>%
  pivot_longer(cols = selected_cols,
               names_to = "cost_type",
               values_to = "cost") %>%
  group_by(outcome_twoarm, cost_type) %>%
  summarise(
    mean_cost = mean(cost, na.rm = TRUE),
    se_cost = sd(cost, na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  ggplot(aes(x = factor(outcome_twoarm), y = mean_cost, 
             fill = factor(outcome_twoarm))) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_cost - se_cost, 
                    ymax = mean_cost + se_cost),
                width = 0.2) +
  facet_wrap(~cost_type, scales = "free_y") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  labs(x = "Study Arm",
       y = "Mean Cost (USh)",
       fill = "Study Arm") +
  theme(legend.position = "bottom")
plot(cost_barplot)

# Linear regression analysis
lm_model <- data %>%
  lm(total_cost ~ factor(outcome_twoarm) + age_cat + sex + 
     far_distance + enroll_period, data = .)

# Quantile regression (median)
qr_model <- data %>%
  rq(total_cost ~ factor(outcome_twoarm) + age_cat + sex + 
     far_distance + enroll_period, tau = 0.5, data = .)

# Create tidy summary of regression results
regression_results <- bind_rows(
  tidy(lm_model) %>% mutate(model = "Linear Regression"),
  tidy(qr_model) %>% mutate(model = "Quantile Regression")
)

# Analyze contribution of each cost component
cost_contribution <- data %>%
  group_by(outcome_twoarm) %>%
  summarise(across(
    selected_cols,
    ~mean(./loss_v2, na.rm = TRUE)
  )) %>%
  pivot_longer(cols = -outcome_twoarm,
               names_to = "cost_component",
               values_to = "proportion")

# Create proportion plot
cost_proportion_plot <- ggplot(cost_contribution, 
       aes(x = factor(outcome_twoarm), 
           y = proportion, 
           fill = cost_component)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(x = "Study Arm",
       y = "Proportion of Total Cost",
       fill = "Cost Component") +
  theme(legend.position = "right")

# List of outputs
list(
  summary_stats = cost_summary,
  box_plot = cost_boxplot,
  bar_plot = cost_barplot,
  proportion_plot = cost_proportion_plot,
  linear_regression = tidy(lm_model),
  quantile_regression = tidy(qr_model)
)
```

```{r}
cost_barplot
```




```{r}
data <- data %>%
  mutate(sex)

sex, age_atenroll


hist(data$sv1_timetraveltypical)
sum(data$sv1_timetraveltypical >= 3)

temp2 <- data %>%
  filter(!is.na(choice_initial)) %>%
  filter(sv1_timetraveltypical >= 3) %>%
  select(sex, age_atenroll, employment_recode, faraway, hourlywage, choice_initial, sv1_timetraveltypical)
```

